PUBPOL 543

Deliverable 1: "Univariate Plot"

Note: The dataset used for this assignment was sourced from the US Government's open data repository: https://catalog.data.gov/ (last accessed, February 5, 2023) 

The dataset contains information on data breach incidents reported by legal entities from different industries. In this, the Washington state laws requires legal entities to notify authorities whenever they have been exposed to data breaches compromising information on more than 500 Washingtonians. 

```{r}
#Providing the link to the data in the Github repository:
location='https://github.com/elenapopic543/Univariate_Deliverable1/raw/main/'
file='Data_Breach_Notifications_Affecting_Washington_Residents.csv'
link=paste0(location,file)

#Reading-in the data from the Github repository location:
data1 <- read.csv(file=url(link), header=TRUE, stringsAsFactors=FALSE)
```

```{r}
#Examining the variables in the dataframe:
names(data1)
```

```{r}
#Examining data types:
str(data1,width = 70,strict.width='cut')
```

```{r}
#Examining raw values:
head(data1$IndustryType,20) 
```


```{r}
#Creating a frequency table:
absoluteT=table(data1$IndustryType,
                exclude = 'nothing')
absoluteT
```

```{r}
#Transforming the counts into proportions:
prop.table(absoluteT)
```

```{r}
#Translating proportions into percents:
propT=prop.table(absoluteT)*100
propT
```

```{r}
#Turning the frequency table into a dataframe:
(tableFreq=as.data.frame(absoluteT))
```

```{r}
#Renaming columns:
names(tableFreq)=c("Industry","Count")
#Adding percents:
tableFreq$Percent=as.vector(propT)
tableFreq
```

```{r}
library(ggplot2)
#Creating the base object:
base = ggplot(data = tableFreq, 
             aes(x = Industry,
                 y = Count))
#Adding layers to create default barplot:
plot1 = base + geom_bar(fill ="gray",
                        stat = 'identity') 
#plot1
```

```{r}
#Adding titles:
titleText='Data Breach Incidents by Industry in Washington State (2016-2022)'
sourceText='Source: DATA.GOV'

plot2 = plot1 + labs(title=titleText,
                     x =NULL, 
                     y = NULL,
                     caption = sourceText) 
#plot2
```

```{r}
#Using percent values instead of counts on the Y-axis:
base= ggplot(data = tableFreq, 
             aes(x = Industry,
                 y = Percent)) 
plot1 = base + geom_bar(fill ="gray",
                        stat = 'identity') 
plot2 = plot1 + labs(title=titleText,
                     x =NULL, 
                     y = NULL,
                     caption = sourceText)
#Customizing the axes:
library(scales) 
plot3 = plot2 + scale_y_continuous(breaks=c(0,10,20,30,40,50),
                                   limits = c(0,50), 
                                   labels=unit_format(suffix = '%')) 
#plot3
```

```{r}
#Adding the actual percent values (rounded to two decimal places):
paste0(round(tableFreq$Percent,2), '%')
```

```{r}
#Generating values for column labels
LABELS = paste0(round(tableFreq$Percent,2), '%')
```

```{r}
#Ordering the bars by height:
tableFreq=tableFreq[order(tableFreq$Percent),]
tableFreq
#Saving the order:
(industriesOrd=tableFreq[order(tableFreq$Percent),'Industry'])
```

```{r}
#Loading the Color Brewer package:
library(RColorBrewer)

#Consolidating the code and generating the final plot: 

# Generating values for column labels
LABELS = paste0(round(tableFreq$Percent,2), '%')

# Specifying the base of the plot
base = ggplot(data = tableFreq, aes(x = Industry, y = Percent)) 

# Ordering the bars by height
base = base + scale_x_discrete(limits=industriesOrd) 

# Changing the theme to classic and adding horizontal lines
base = base + theme_classic() + theme(panel.grid.major.y = element_line(colour = "gray80", linetype = "dashed"))

# Adding layers to the base and defining the color scheme
 plot1 = base + geom_bar(fill = brewer.pal(5,"Blues")[3], stat = 'identity')
 
# Adding title and caption
plot2 = plot1 + labs(title = titleText, x = "Industry Affected", y = "Percent of Data Breaches", caption = sourceText)

# Customizing and labeling the values on the Y-axis with a % suffix:
plot3 = plot2 + scale_y_continuous (breaks = c(0, 10, 20, 30, 40, 50), limits = c(0, 50), labels = unit_format(suffix = '%')) 

# Moving the plot title to the center, and caption to left, and adjusting spacing:
plot4 = plot3 + theme(plot.caption = element_text(hjust = 0, margin = margin(t = 5), size=9), 
                    plot.title = element_text(hjust = 0.5, margin = margin(b = 15), size = 13),
                      axis.title.x = element_text(margin = margin(t = 8), size=10), 
                      axis.title.y = element_text(margin = margin(r = 8), size=10))

# Adding column labels and adjusting size/location
plot5 = plot4 + geom_text(vjust = -0.1, size = 3, aes(y = Percent, label = LABELS))

# Displaying the final plot
plot5
```
